---
title: "Homework Assignment 1"
author: "Ji Wei Mok"
date: 'Assigned: Oct 24, 2020, Due Sun Nov 01, 2020 11:59PM'
output:
  #pdf_document: default
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Data frame basics
## Reading my data

```{r, cache=TRUE}
nyc <- read.csv("nycflights.csv")

# Quick exploration of the data
str(nyc)
summary(nyc)

```

## Cleaning my data

```{r}
# Deleting the column X
nyc$X <- NULL

# Factorize columns that need to be a factor
nyc$origin <- as.factor(nyc$origin)
nyc$tailnum <- as.factor(nyc$tailnum)
nyc$month <- as.factor(nyc$month)
nyc$dest <- as.factor(nyc$dest)
nyc$carrier <- as.factor(nyc$carrier)
nyc$flight <- as.factor(nyc$flight)

# Re-check my data
str(nyc)
summary(nyc)
```

# Data Exploration
```{r}
# Number of airlines
numairlines <- nlevels(nyc$carrier)
numairlines

# Number of flights with airline code OO

numairlineOO <- nrow(nyc[nyc$carrier=="OO", ])
numairlineOO

# How long is the shortest flight out of NYC airport
shortestflight <- min(nyc$air_time, na.rm = TRUE)

# How many UA flights on Jan 12th 2013

numUA <- nrow(nyc[nyc$carrier == "UA" & nyc$month == "1" & nyc$day == "13", ])
numUA

```
## Arrival Delay
```{r}
# Average arrival delay for all airports and airlines combined in Jan 2013
avgdelay <- mean(nyc$arr_delay[nyc$month == "1" & nyc$year=="2013"], na.rm = TRUE)
avgdelay

# Median arrival delay for all airports and all airlines combined in Jan 2013?
mediandelay <- median(nyc$arr_delay[nyc$month == "1" & nyc$year=="2013"], na.rm = TRUE)
mediandelay

```
The distribution of arrival delays is right skew since median is less than mean

## Airline Performance
```{r}
# Average arrival delays by airlines
avgarrdelayairline <- tapply(nyc$arr_delay, nyc$carrier, mean, na.rm=TRUE)
avgarrdelayairline

# Bar Plot of average arrival delays for all airlines
barplot(avgarrdelayairline)

```
Airline with highest average arrival delay is `r names(avgarrdelayairline[avgarrdelayairline==max(avgarrdelayairline)])`
Airline with smallest average arrival delay is `r names(avgarrdelayairline[avgarrdelayairline==min(avgarrdelayairline)])`
Airline with negative average arrival delay is `r names(avgarrdelayairline[avgarrdelayairline<0])`

## Air Gain
```{r}
# Average airgain
nyc$airgain <- nyc$dep_delay - nyc$arr_delay

# average airgain for different airlines
avgairgainairlines <- tapply(nyc$airgain, nyc$carrier, mean, na.rm=TRUE)
barplot(avgairgainairlines)
```
Airlines gain `r mean(nyc$airgain, na.rm=TRUE)` when in air
Airline `r names(avgairgainairlines[avgairgainairlines==max(avgairgainairlines)])` does best job
Airline `r names(avgairgainairlines[avgairgainairlines==min(avgairgainairlines)])` does worst job

## Reading new data
```{r, cache=TRUE}
planes <- read.csv("planes.csv")

# Quick exploration of the data
str(planes)
summary(planes)
```
## Cleaning new data

```{r}
# Deleting the column X
planes$X <- NULL

# Factorize columns that need to be a factor
planes$type <- as.factor(planes$type)
planes$tailnum <- as.factor(planes$tailnum)
planes$manufacturer <- as.factor(planes$manufacturer)
planes$model <- as.factor(planes$model)
planes$engine <- as.factor(planes$engine)
planes$seats <- as.factor(planes$seats)

# Re-check my data
str(planes)
summary(planes)

```

## Merging Data Frames

```{r, cache=TRUE}
joined <- merge(nyc,planes,by.x="tailnum",by.y="tailnum",all.x = TRUE, all.y=TRUE)
# Re-check my data
str(joined)
summary(joined)

```
## Questions

```{r}
#What are top 10 manufacturers of planes that fly out of NYC airports?
topmanufacturers <- names(head(sort(table(joined$manufacturer),decreasing=TRUE),n=10))
topmanufacturers

#Average arrival delays by manufacturers?
avgarrdelaymanufacturer <- tapply(joined$arr_delay,joined$manufacturer, mean, na.rm=TRUE)
avgarrdelaymanufacturer

#What is the most popular seat configuration of UA planes flying out of NYC?
UAdata <- joined[joined$carrier=="UA", ]
topseat <- names(head(sort(table(UAdata$seats),decreasing=TRUE),n=1))
topseat

#What is the average age of the planes in year 2020?
avgage <- 2020-mean(planes$year,na.rm=TRUE)
avgage

#Which aircraft has flown the most?
freq <- table(joined$tailnum)
maxFreq <- which.max(freq)
names(maxFreq)

```